The GitHub Open Source Development Process \author{Kevin Peterson\\ Department of Information Technology\\ Clinical Informatics Support Systems\\ Mayo Clinic\\ 200 1st Street\\ Rochester, MN 55905\\ \small \texttt{peterson.kevin@mayo.edu} }

### 2013-04-20 paper/main.tex
Updated the conclusion.

As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification~\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes. Elicitation, Analysis and Negotiation were natural fits for the issue tracker. This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issue tracker itself allowed for dialogs in the form of comments. Because customer involvement is critical to these activities~\cite{paetsch2003requirements}, ease of contribution was a main concern. OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
Validation and Management are areas of requirements engineering that GitHub has interesting solutions for. First, in order to track (or validate) changes to the specification, when the actual specification is modified, the commit that modified it can be traced to a named issue by simply putting the issue number in the commit log. This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue. As the normal OMG\textsuperscript{\textregistered} process proceeds, requirements can be managed through ``tagging'' the issues. By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own work flow. For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration

@@

\section{Conclusion}
The data suggests an \textit{evolution} in process, but not necessarily a large shift. A small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure~\ref{eq:variance}, and the distribution of committers per repository in Figure~\ref{fig:committers_histogram}. Figure~\ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories. This may indicate a lowered entry barrier for participation, and is consistent with the conclusions reached in Research Question 2.

The data does not reflect some traditional aspects of OSS development. Notably, it was not observed that there were far more \emph{Issue Reporters} than code \emph{Committers}, which is contrary to what the Apache Server report concluded~\cite{mockus2000case}. A possible explanation is the amount of variability in the data collected. Assertion~\ref{eq:variance}, which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables. This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises. Given this, we cannot state any conclusions regarding participation rates of the \emph{Issue Reporters}. More research into categorizing GitHub repositories, and the role of \emph{Issue Reporters} in them, is needed.

The most notable \textit{evolution} in OSS process driven by GitHub, we conclude, is the lowered barrier to participation. Qualitative analysis conducted in Research Question 2 suggest a cultural shift~\cite{begel2013social,github_interview} in the participation process. The data also supports high levels of participation, especially Figure~\ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code. This, we conclude, represents an important \textit{evolution} in OSS development process.

\section{Acknowledgments}
We thank Gilberto Fragoso, Ph.D., Paul Alexander, John Britton, and all of the GitHub Staff for interviews, insights and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.

>>>>

As a case study, the CTS2 OMG\textsuperscript{\textregistered} Specification~\cite{cts2} used a GitHub issue tracker\footnote{https://github.com/cts2/cts2-specification/issues} to track specification changes. Elicitation, Analysis and Negotiation were natural fits for the issue tracker. This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issue tracker itself allowed for dialogs in the form of comments. Because customer involvement is critical to these activities~\cite{paetsch2003requirements}, ease of contribution was a main concern. OMG\textsuperscript{\textregistered} has a predescribed format for requirements Documentation, but using the GitHub issue API,\footnote{http://developer.github.com/v3/issues/} much of this documentation can be automated.
Validation and Management are areas of requirements engineering that GitHub has interesting solutions for. First, in order to track (or validate) changes to the specification, when the actual specification is modified, the commit that modified it can be traced to a named issue by simply putting the issue number in the commit log. This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue. As the normal OMG\textsuperscript{\textregistered} process proceeds, requirements can be managed through ``tagging'' the issues. By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own work flow. For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration

@@

\section{Conclusion}
The analysis suggests an \textit{evolution} in process, but not necessarily a large shift. We classify this as an \textit{evolution} because although many of the traditional aspects of the OSS development process remain, there are notable differences. For example, a small subset of core developers, a traditional aspect of OSS development~\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}, seems supported by both the summary data in Figure~\ref{eq:variance}, and the distribution of committers per repository in Figure~\ref{fig:committers_histogram}. Figure~\ref{fig:committers_percentage_pie_chart} shows a significant number of participators, albeit with small contribution amounts, in many repositories. This may indicate a lowered entry barrier for participation, and is consistent with the conclusions reached in Research Questions 1 and 2.

The data does not, however, reflect some other traditional aspects of OSS development. Notably, it was observed that data gathered for Hypothesis 2 and Hypothesis 3 seemed contrary to what the Apache Server report concluded~\cite{mockus2000case}. A possible explanation is the amount of variability in the data collected. Assertion~\ref{eq:variance}, which is derived from the summary data shown in Table~\ref{table:summary_stats}, suggests large amounts of variety in the observed variables. This may suggest a wide range of repository purposes, or specifically, that not all repositories on GitHub are intended to be consumable software artifacts for users, or ``customers.'' A large number of repositories may not be intended for customer consumption, but merely as example code, experimentation, or exercises. Given this, and specific reasons outlined in the hypotheses themselves, we cannot state any definitive conclusions regarding these two hypotheses. More research into categorizing GitHub repositories is needed.

The most notable \textit{difference} in OSS process driven by GitHub, considering both the quantitative and qualitative analysis, is the lowered barrier to participation. Qualitative analysis conducted in Research Question 2 suggests a cultural shift~\cite{begel2013social,github_interview} in the participation process. The data also supports high levels of participation, especially Figure~\ref{fig:committers_percentage_pie_chart}, which shows that many repositories have large numbers of developers committing small amounts of code. This, we conclude, represents an important \textit{evolution} in the OSS development process.

\section{Acknowledgments}
We thank Gilberto Fragoso, Ph.D. (NIH/NCI), Paul Alexander (NCBO), John Britton (GitHub), and all of the GitHub Staff for interviews, insights and ideas, Harold Solbrig, Craig Stancl, and Dick Hedger for guidance, and Rick Kiefer for his review.
